### 2018-02-12 Writing/Thesis/content/chapter-conclusion.tex
more writing

% Summary of findings and limitations: what has been covered LIMITATION: SMALL DATA SETS ESP!!
The evaluation showed a significant advantage regarding the Kendall rank correlation for all tested methods in comparison to the best algorithm baseline, except instance based label ranking in the standard configuration and label ranking by pairwise comparison for the full meta data. However, regarding the practically possibly more important measure of best three loss, only the three regression based variants with linear regression, M5P and random forest as the regression learners can hold a significant advantage. When evaluating best three loss on reduced meta data that does not incorporate probing, although the evaluated data indicates a slight advantage of those same regression models over the baseline still, it is not statistically significant anymore. Furthermore, some label ranking methods even show a significant disadvantage in comparison to the baseline for the best three loss and loss for both evaluation on the full and reduced meta data. Thus it can be included that a regression based approach seems to be generally more promising than the preference based approach, but is not better than the baseline regarding the best three loss when evalluated on the selected meta data without probing.

One of the main limitations of the findings in this thesis is that the data sets used in the evaluation process where not very large, being limited to a maximum of 100 features and 1000 instances. Often, data sets can get much larger and the question whether this ranking approach would then still pose an efficient and, maybe even more importantly accurate enough solution for predicting classifiers remains.

% Practical applications / limitations: assess value / relevance / implications: What does it mean for theory, what for practice
On the basis of these results, it can be concluded that a causal connection exists between certain meta features of a data set and the predictive accuracy of classification algorithms for this data set, which can be exploited to a degree by regression models and label ranking models to predict a ranking of classification algorithms. A practical application of these findings, apart from the direct use as a ranker, may be to incorporate the implementation or parts of it in another Auto-ML tool as a search heuristic, similar to how some Auto-ML solutions like AUTO-SKLEARN already benefits from meta learning \cite{feurer2015efficient}. However, some additional work may have to be done in extension to this thesis in order for a sensible integration.

@@

% Use this to predict other measure
In a similar manner to extending the algorithms for which a performance measure is predicted, one could use the tool to predict different performance measures. For this, the tool does not even need to be adapted, as it is agnostic to the semantics behind the measure it predicts, it only needs to be numeric. One possibility for this would be the prediction of build or prediction times of learning algorithms, or a measure that combines time and accuracy \cite{DBLP:journals/ml/AbdulrahmanBRV18}.

@@

% Additional baseline
Last, it could be interesting to implement an additional baseline that averages ranks instead of counting how many times an algorithm is ranked first. This would help to determine if the surprising ranking returned by the best algorithm baseline (see Fig. \ref{tab:bestAlgorithmRanking}
) still holds when using a different baseline. Naturally this new baseline then could also be used to re-evaluate the implemented ranking concepts.

>>>>

% Summary of findings and limitations: what has been covered LIMITATION: SMALL DATA SETS ESP!!
The evaluation showed a significant advantage regarding the Kendall rank correlation for all tested methods in comparison to the best algorithm baseline, except instance based label ranking in the standard configuration and label ranking by pairwise comparison for the full meta data. However, regarding the practically possibly more important measure of best three loss, only the three regression based variants with linear regression, M5P and random forest as the regression learners can hold a significant advantage. When evaluating best three loss on reduced meta data that does not incorporate probing, although the evaluated data indicates a slight advantage of those same regression models over the baseline still, it is not statistically significant anymore. Furthermore, some label ranking methods even show a significant disadvantage in comparison to the baseline for the best three loss and loss for both evaluation on the full and reduced meta data. Thus it can be included that a regression based approach seems to be generally more promising than a preference based approach, but is not better than the baseline regarding the best three loss when evaluated on the selected meta data without probing. Out of the regression based variants, random forest seems to be the most promising one.

The main limitation of the findings in this thesis is that the data sets used in the evaluation process where not very large, being limited to a maximum of 100 features and 1000 instances. Often, data sets can get much larger and the question whether this ranking approach would then still pose an efficient regarding calculation of meta features and prediction times and, maybe even more importantly, an accurate enough solution for predicting classifiers remains.

% Practical applications / limitations: assess value / relevance / implications: What does it mean for theory, what for practice
On the basis of these results, it can be concluded that a there is an indication of a connection between certain meta features of a data set and the predictive accuracy of classification algorithms for this data set, which can be exploited to a degree by regression models and label ranking models to predict a ranking of classification algorithms. A practical application of these findings, apart from the direct use as a ranker, may be to incorporate the implementation or parts of it in another Auto-ML tool as a search heuristic, similar to how some Auto-ML solutions like AUTO-SKLEARN already benefits from meta learning \cite{feurer2015efficient}. However, some additional work may have to be done in extension to this thesis in order for a sensible integration.

@@

% Use this to predict other measure
In a similar manner to extending the algorithms for which a performance measure is predicted, one could use the tool to predict different performance measures. For this, the tool does not even need to be adapted, as it is agnostic to the semantics behind the measure it predicts, it only needs to be numeric. One possibility for this would be the prediction of build or prediction times of learning algorithms, or a measure that combines time and accuracy like A3R \cite{DBLP:journals/ml/AbdulrahmanBRV18}.

@@

% Additional baseline
Last, it could be interesting to implement an additional baseline that averages ranks instead of counting how many times an algorithm is ranked first. This would help to determine if the ranking returned by the best algorithm baseline (see Fig. \ref{tab:bestAlgorithmRanking}
) still holds when using a different baseline. Naturally this new baseline then could also be used to re-evaluate the implemented ranking concepts.

### 2018-02-12 Writing/Thesis/content/chapter-evaluation.tex
more writing

% Evaluation Measures used & computed
As rankings returned by this oracle represent the ideal solution, the predicted rankings are compared with them through the measures presented in chapter \ref{chapter:fundamentals}. These include the Kendall rank correlation coefficient, loss and best three loss. For all three measures, it is computed if any advantages or disadvantages one solutions has is significant, by means of the Mann-Whitney U. The regression based rankers pose a special case: internally, they predict a performance value for each classifier, which therefore can be compared to the actual performance values to show how well they are predicted. This is done by computing the root mean square error between the predicted and actual performance values. In summary, it is desirable for the rankers to come as close as possible to the correct ranking, which is reflected by a high rank correlation and a low loss and root mean square error.

% Baseline
Furthermore, we are interested in the question as to what degree knowing about the properties of a data set influences the quality of rankings. Therefore, a ranker that is agnostic of the meta features of a query data set, that is that will always return the same ranking for any data set, is used as a baseline. More specifically, a best algorithm strategy that iteratively determines a ranking by counting the number of data sets where an algorithm is the best choice, is implemented. That means the first algorithm in the best algorithm ranking is the classifier that performs best on most data sets, the second is the best on most data sets when only rankings excluding the first algorithm are considered, and so forth. This baseline is evaluated in the same way as the preference rankers; as it does not predict performance values, the root mean square error of predicted performance values cannot bet computed. Ideally, rankings returned by the preference and regression based rankers should then be statistically significant better than the baseline, which will be determined by calculating the Mann-Whitney U and determining whether its P-value.

@@

% Which data sets
Data sets for the analysis where gathered from OpenML. From all data sets considered as 'active' on OpenML, which yields 2050 data sets that have been approved and with no severe issues found for them so far \cite{openMLGuide}, all data sets that comply with the constraints of the learning problem at hand where selected. Only data sets with a defined, nominal target feature that are in the .ARFF format and where not specifically designed for streaming analysis\footnotemark{} where considered, which resulted in a reduced selection of 812 instances. Since this amount of data sets hindered the evaluation considerably, large data sets with more than 1000 instances or 100 features where removed as well, leading to a final selection of 448 data sets. Even though evaluation was only carried out on the shorter list, both the list of 812 and 448 data sets are included in the supplementary material for completeness.

\footnotetext{Data sets that contained the substring 'BNG' for Bayesian Network Generator, which contain data artificially generated by a Bayesian Network\cite{van2014algorithm} for the sake of data stream analysis where not included in the evaluation.}

% Which meta features
As for the meta features, all meta features discussed in chapter \ref{sec:approach} where used. Initially, it was intended to download the computed meta features from OpenML, but since missing values for meta features ranged from roughly 25\% up to 89\% (except for a few simple meta features which were included for all data sets) it was decided to compute them anew. Meta features where calculated with the help of an implementation provided by OpenML \cite{openMLEvaluationEngine}.Two data sets, one including the downloaded meta features (for all 812 data sets) and one including the computed meta features (for the reduced set of 448 data sets) are thus included in the supplementary material.

@@

% How rest of the evaluation carried out
All other experiments where carried out on a Windows Machine running Windows 10 Education (version 1709) with a Intel(R) Core(TM) i5-4200U CPU running at 1.60GHz (2.30GHz max) and 12 GB RAM. It is notable that especially the time-related results are to be viewed in regard to this setup. No timeouts where used in this evaluation, and the rankers where evaluated on the training data by means of leave-one-out estimation, that is for the n data sets for which performance values where recorded, they where trained with the meta data - performance examples for n - 1 data sets, and queried for the remaining data set. As n = 448 in this case, the detailed values for each data set are not included in the discussion of the results. They can, however, be found in the supplementary material, offering clues to how each ranker performed in the prediction of a ranking for each of the used data sets, identified by their ID on OpenML.

@@

- on the full meta data set, all approaches except for instance based with kemeny young and pairwise comparison

@@

Because the ranking itself is implemented as a means of speeding up the process of algorithm selection, the time it takes to rank is not negligible. The full results of how long it takes to build the respective ranking models and how much time they need for predictions can be taken from Table \ref{tab:times}. Results are rounded to the full millisecond, as due to the times being measured with a stop watch in the code, more accurate measurements are not possible anyways. The times for building the regression rankers are not surprising; the regression models take far more time to be built than most of the label ranking methods, most likely due to the fact the regression based rankers have to train 22 regression models in order to be trained themselves. However, it is notable that while the ranking by Pairwise Comparison was not among the best solutions concerning the accuracy of the results, it is the ranker with the highest mean build time for full meta data, and second highest for a reduced meta data set, whereas in other cases the high quality of results is correlated with a high time invested in the building of the ranker likewise. The Random Forest ranker, for example, is the ranker with the highest build time, and also with the rankings of the generally highest quality. But the most important fact to be taken away from Table \ref{tab:times} is that the prediction times for all rankers are short enough to be negligible, with the mean prediction time never being higher than five milliseconds and the maximum never exceeding 50 milliseconds. Furthermore, while the build times are higher, these are not as important due to the fact that each model only needs to be trained once. Although in a real-world scenario, calculation times for meta features also have to be taken into account when ranking, which therefore will be discussed in the following paragraph.

@@

Table \ref{tab:metaFeatureTimes} shows the times required For the full meta features, on average approximately 135 milliseconds would be added to the prediction times 3 milliseconds whithout probing

@@

Further insights to be noted are the accuracy of predictions of performance values for classifiers by the regression models, and the static ranking constructed by the best algorithm baseline. The accuracy of the predictions achieved by the regression based rankers are depicted in Table \ref{tab:rootMeanSquareError}, for full and reduced meta data respectively. The first thing that attracts attention in this table is the absurdly high values for Linear Regression and M5P, however, these are all caused by the very high maximum value on one data set with id 685. If these extreme maxima remain out of consideration for Linear Regression and M5P (the other algorithms do not have this problem) more reasonable values are obtained. For M5P, on the full meta data the new maximum is 311.526, mean 8.127 and standard deviation 18.407; on the reduced meta data it is a maximum of 474.692, mean 15.444 and standard deviation 27.433. For Linear Regression, on the full meta data the new maximum is 142.996, mean 9.281 and standard deviation 8.438; on the reduced meta data it is a maximum of 131.456, mean 14.351 and standard deviation 10.299. However, these values are still worse than the ones obtained for the Random Forest and REPTree variants. It comes as little surprise that the Random Forest Ranker also delivers the most accurate results for performance estimates, but as these best values are more than 6\% off on average for the full meta data and even more than 10\% off on average for the reduced meta data, it is questionable whether these prediction can be useful.

>>>>

% Evaluation Measures used & computed
As rankings returned by this oracle represent the ideal solution, the predicted rankings are compared with them through the measures presented in chapter \ref{chapter:fundamentals}. These include the Kendall rank correlation coefficient, loss and best three loss. For all three measures, it is computed if any advantage or disadvantage one solutions has is significant, by means of the Mann-Whitney U and its P-value. The regression based rankers pose a special case: internally, they predict a performance value for each classifier, which therefore can be compared to the actual performance values to show how well they are predicted. This is done by computing the root mean square error between the predicted and actual performance values. In summary, it is desirable for the rankers to come as close as possible to the correct ranking, which is reflected by a high rank correlation and a low loss and root mean square error.

% Baseline
Furthermore, we are interested in the question as to what degree knowing about the properties of a data set influences the quality of rankings. Therefore, a ranker that is agnostic of the meta features of a query data set, that is that will always return the same ranking for any data set, is used as a baseline. More specifically, a best algorithm strategy that iteratively determines a ranking by counting the number of data sets where an algorithm is the best choice, is implemented. That means the first algorithm in the best algorithm ranking is the classifier that performs best on most data sets, the second is the best on most data sets when only rankings excluding the first algorithm are considered, and so forth. This baseline is evaluated in the same way as the preference rankers; as it does not predict performance values, the root mean square error of predicted performance values cannot bet computed. Ideally, rankings returned by the preference and regression based rankers should then be statistically significant better than the baseline.

@@

% Which data sets
Data sets for the analysis where gathered from OpenML. From all data sets considered as 'active' on OpenML, which yields 2050 data sets that have been approved with no severe issues found for them so far \cite{openMLGuide}, all data sets that comply with the constraints of the learning problem at hand where selected. Only data sets with a defined, nominal target feature that are in the .ARFF format which where not specifically designed for streaming analysis\footnotemark{} where considered, which resulted in a reduced selection of 812 instances. Since this amount of data sets hindered the evaluation considerably, large data sets with more than 1000 instances or 100 features where removed as well, leading to a final selection of 448 data sets. Even though evaluation was only carried out on the shorter list, both the list of 812 and 448 data sets are included in the supplementary material for completeness.

\footnotetext{Data sets that contained the substring 'BNG' in their name for Bayesian Network Generator, which contain data artificially generated by a Bayesian Network \cite{van2014algorithm} for the sake of data stream analysis where not included in the evaluation.}

% Which meta features
As for the meta features, all meta features discussed in chapter \ref{sec:approach} where used. Initially, it was intended to download the computed meta features from OpenML, but since missing values for meta features ranged from roughly 25\% up to 89\% (except for a few simple meta features which were included for all data sets) it was decided to compute them anew. Meta features where calculated with the help of an implementation provided by OpenML \cite{openMLEvaluationEngine}. Two data sets, one including the downloaded meta features (for all 812 data sets) and one including the computed meta features (for the reduced set of 448 data sets) are thus included in the supplementary material. Since the meta data also includes probing, which is generally an expensive operation and furthermore does not aid in proving the point of a connection of properties of a data set and the performance of classifiers, additionally to the full meta features all evaluation measures where computed without using those meta features that use probing.

@@

% How rest of the evaluation carried out
All other experiments where carried out on a Windows Machine running Windows 10 Education (version 1709) with a Intel(R) Core(TM) i5-4200U CPU running at 1.60GHz (2.30GHz max) with 12 GB RAM. It is notable that especially the time-related results are to be viewed in regard to this setup. No timeouts where used in this evaluation, and the rankers where evaluated on the training data by means of leave-one-out estimation, that is for the n data sets for which performance values where recorded, they where trained with the meta data - performance examples for n - 1 data sets, and queried for the remaining data set. As n = 448 in this case, the detailed values for each data set are not included in the discussion of the results. They can, however, be found in the supplementary material, offering clues to how each ranker performed in the prediction of a ranking for each of the used data sets, identified by their ID on OpenML.

@@

When looking at the results depicted in Table \ref{tab:evaluationResults}, it must first be noted that positive findings regarding the Kendall rank correlation can be reported. For the full set of meta data, six out of the eight implemented alternatives outperform the baseline, all of them significantly, as can be learned from Table \ref{tab:significanceResults} which shows the results of the significance tests, with random forest having the highest correlation of 0.495. In contrast, the two approaches that did not overtake the baseline, instance based label ranking in the default configuration and label ranking by pairwise comparison, show a significant disadvantage compared to the baseline, and thus also compared to the other ranking approaches. Regarding loss, only linear regression and random forest show a significant advantage, with all label ranking approaches except for instance based label ranking having a significant disadvantage in contrast. Moving on to best three loss, the picture largely remains the same, except that surprisingly, REPTree now also has a significant advantage over the baseline. The lowest value here is achieved by random forest with a best three loss of 1.308, in comparison to 2.440 for the best algorithm baseline. These values are illustrated in Figure \ref{fig:lossGraphics}, with a scatter plot for individual best three loss values on the data sets.

Compared to the evaluation on the full set of meta data, the results on the reduced set without out probing are less distinctive. For the Kendall Rank correlation, linear regression, M5P and random forest are significantly better than the baseline, and for label ranking, both of the instance based approaches with non-default parameters as well, while the other label ranking approaches are significantly worse than the baseline. In total, random forest is still the best alternative with a correlation of 0.439, slightly worse than the correlation achieved on the full meta data. For both loss measures however, while the disadvantages of the label ranking approaches remain, none of the eight alternatives achieves a significantly better loss or best three loss than the baseline anymore.

Summarizing the observed results, in general, a regression based approach shows an indication to be stronger than a preference based approach. Specifically, using random forest as a regression model is the best solution observed in the evaluation, and label ranking by pairwise comparison the weakest, as it falls behind the other ranking approaches and the baseline often.

@@

Because the ranking itself is implemented as a means of speeding up the process of algorithm selection, the time it takes to rank is not negligible. The full results of how long it takes to build the respective ranking models and how much time they need for predictions can be taken from Table \ref{tab:times}. Results are rounded to the full millisecond, as due to the times being measured with a stop watch in the code, more accurate measurements are not possible. The times for building the regression rankers are not surprising; the regression models take far more time to be built than most of the label ranking methods, most likely due to the fact the regression based rankers have to train 22 regression models in order to be trained themselves. However, it is notable that while the ranking by Pairwise Comparison was not among the best solutions concerning the accuracy of the results, it is the ranker with the highest mean build time for full meta data, and second highest for a reduced meta data set, whereas in other cases the high quality of results is correlated with a high time invested in the building of the ranker. The Random Forest ranker, for example, is the ranker with the highest build time, and also with the rankings of the generally highest quality. But the most important fact to be taken away from Table \ref{tab:times} is that the prediction times for all rankers are short enough to be negligible, with the mean prediction time never being higher than five milliseconds and the maximum never exceeding 50 milliseconds. Furthermore, while the build times are higher, these are not as important due to the fact that each model only needs to be trained once. Although in a real-world scenario, calculation times for meta features also have to be taken into account when ranking, which therefore will be discussed in the following paragraph.

@@

Table \ref{tab:metaFeatureTimes} shows the times required for the computation of groups of meta features as discussed in chapter \ref{sec:approach}. As can be taken from the table, the computation of the full meta features takes approximately 135 milliseconds on average, while the simple meta features without probing add approximately 3 milliseconds to the prediction times on average. Regarding these computation times it is also notable that among the probing meta feature groups, some are considerably more expensive than others, so that in the future it might be sensible to consider only light probing with probing meta feature groups that are not as expensive.

@@

Further insights to be noted are the accuracy of predictions of performance values for classifiers by the regression models, and the static ranking constructed by the best algorithm baseline. The accuracy of the predictions achieved by the regression based rankers are depicted in Table \ref{tab:rootMeanSquareError}, for full and reduced meta data respectively. The first thing that attracts attention in this table is the absurdly high values for Linear Regression and M5P, however, these are all caused by the very high maximum value on one data set with id 685. If these extreme maxima remain out of consideration for Linear Regression and M5P (the other algorithms do not have this problem) more reasonable values are obtained. For M5P, on the full meta data the new maximum is 311.526, mean 8.127 and standard deviation 18.407; on the reduced meta data it is a maximum of 474.692, mean 15.444 and standard deviation 27.433. For Linear Regression, on the full meta data the new maximum is 142.996, mean 9.281 and standard deviation 8.438; on the reduced meta data it is a maximum of 131.456, mean 14.351 and standard deviation 10.299. However, these values are still worse than the ones obtained for the Random Forest and REPTree variants. It comes as little surprise that the Random Forest Ranker also delivers the most accurate results for performance estimates, but as these best values are more than 6\% off on average for the full meta data and even more than 10\% off on average for the reduced meta data, it is questionable whether these prediction can be useful. The usefulness is dependent on the task these predictions would be used for an whether if requires very accurate predictions of performance values.

### 2018-02-12 Writing/Thesis/content/chapter-related-work.tex
more writing

\section{Predicting Rankings}
%% Ranking
% Incorporating times
In contrast to ranking solely based on classifier performances, \citeauthor{DBLP:journals/ml/AbdulrahmanBRV18} investigated an approach to extend existing ranking methods by incorporating the time needed for the evaluation of the classifier \cite{DBLP:journals/ml/AbdulrahmanBRV18}. They call this combined measure of accuracy and time A3R, and integrate it into to different ranking approaches. In both cases, this lead to a better mean interval loss, which is a measure that expresses the mean loss experienced in respect to the time that has been allowed for the construction of the ranking.

>>>>

\section{Predicting Rankings}
%% Ranking
% Incorporating times
In contrast to ranking solely based on classifier performances, \citeauthor{DBLP:journals/ml/AbdulrahmanBRV18} investigated an approach to extend existing ranking methods by incorporating the time needed for the evaluation of the classifiers \cite{DBLP:journals/ml/AbdulrahmanBRV18}. They call this combined measure of accuracy and time A3R, and integrate it into two different ranking approaches. In both cases, the authors find that this leads to a better mean interval loss, which is a measure that expresses the mean loss experienced in respect to the time that has been allowed for the construction of the ranking, that is time taken to evaluate top-ranked classifiers.
